Fast and optimal centroiding of faint stars

### 2015-11-30 ms/figure.tex
updated the text

Derivation of an explicit expression for the Fisher matrix $\mathcal{F}$ requires specifying a PSF model.
We use the Moffat profile \citep{moffat} for our PSF simulations. 
Moffat profile is an analytic model for stellar PSFs. It has broader wings than
a simple Gaussian profile. The surface brightness of the Moffat profile is given by
\beq
I(r) = \frac{F(\beta -1)}{\pi \alpha^{2}}[1+(r/\alpha)^{2}]^{-\beta},
\label{mof}
\eeq
where $F$ is the total flux, $\beta$ is a dimensionless parameter, and $\alpha$ is
the length scale of the Moffat profile, with FWHM (hereafter denoted by $\gamma$)
being $2\alpha\sqrt{2^{1/\beta}-1}$. At a fixed $\gamma$, Moffat profiles with lower values
of $\beta$ have broader tails. It is also important to note that for sufficiently large values of the 
parameter $\beta$, the Moffat PSF becomes arbitrarily close to a simple Gaussian PSF.

@@

Throughout this investigation, $\beta$ is held fixed at the fiducial value of 2.5, where SNR
is given by the following expression
\beq
\text{SNR} \simeq \frac{0.478}{\sigma \gamma}.
\eeq

@@

It is worth noting that for any PSF model whose radial light profile is some function of 
$r/\gamma$, CRLB has the same functional form, in that it is proportional to the ratio
between $\gamma$ and SNR. 
For PSF profiles with shorter tails (\eg Gaussian), the prefactor of 0.685 in (\ref{crlbmoffat})
becomes smaller. In particular case of Gaussian PSF, the prefactor is approximately 0.6.

We examine fitting an exact
 PSF profile to the stars. That is, in our Cram\'{e}r-Rao bound saturation tests, we find the best
 estimates of flux and centroid by optimizing the $\chi^{2}$.
 We expect this method to perform better in determing the centroids
 of stars, and deliver RMSE very close to Cram\'{e}r-Rao bound. In the next section,
 we briefly discuss the centroiding methods used in our Cram\'{e}r-Rao bound saturating tests.

\section{Centroiding methods}\label{sec:method}
\begin{description}
%\item{{\bf :N}} \quad \\ 
\item{{\bf Matched filter polynomial centroiding}} \quad Let us consider the case 
in which we have a good estimate of the pixel-convolved PSF at
the position of the faint star under consideration. 
We can smooth the image of the star, by correlating it with the 
full PSF $\mathcal{P}$ at the position of the star.
\begin{eqnarray}
Y^{(s)} &=& Y \star \mathcal{P}, \\
Y^{(s)}_{[i,j]} &=& \sum_{k,l}Y_{[i-k,j-l]}\mathcal{P}_{[k,l]},
\end{eqnarray}
where $Y$ is the image of the star, and $Y^{(s)}$ is called matched filter. 
Matched filter is a model where the data $Y$ is correlated (convolved in the 
case of symmetrical PSF) with the PSF $\mathcal{P}$. It provides an optimal an 
optimal detection map where the peak of the map is the likely position of the 
point source (Lang \emph{et al.}, in preparation).

@@

When the curvature matrix $D$ has a zero (or very close to zero) deteminant,
centroid estimates obtained from equation (\ref{center}) can become arbitrarily 
large which leads to catastrophic outliers. 
In order to tackle this issue, we add a soft regularization term
proportional to $\sigma^{2}$ to the determinant of $D$.

>>>>

Derivation of an explicit expression for the Fisher matrix $\mathcal{F}$ requires specifying a presumed correct 
PSF model.
We use the Moffat profile \citep{moffat} for our PSF simulations. 
Moffat profile is an analytic model for stellar PSFs. It has broader wings than
a simple Gaussian profile. The surface brightness of the Moffat profile is given by
\beq
I(r) = \frac{F(\beta -1)}{\pi \alpha^{2}}[1+(r/\alpha)^{2}]^{-\beta},
\label{mof}
\eeq
where $F$ is the total flux, $\beta$ is a dimensionless parameter, and $\alpha$ is
the scale radius of the Moffat profile, with FWHM (hereafter denoted by $\gamma$)
being $2\alpha\sqrt{2^{1/\beta}-1}$. At a fixed $\gamma$, Moffat profiles with lower values
of $\beta$ have broader tails. It is also important to note that for sufficiently large values of the 
parameter $\beta$, the Moffat PSF becomes arbitrarily close to a simple Gaussian PSF.

@@

Throughout this investigation, $\beta$ is held fixed at the fiducial value of $\beta=$ 2.5, where SNR
is given by the following expression
\beq
\text{SNR} \simeq \frac{0.478}{\sigma \gamma} & \text{   for } \beta = 2.5.
\eeq

@@

It is worth noting that for any PSF model whose radial light profile is some function of 
$r/\gamma$, CRLB has the same functional form, in that it is proportional to the ratio
between $\gamma$ and SNR. 
For PSF profiles with shorter tails (\eg, Gaussian), the prefactor of 0.685 in (\ref{crlbmoffat})
becomes smaller. In particular case of Gaussian PSF, the prefactor is approximately 0.6.

We examine fitting an exact
 PSF profile to the stars. That is, in our Cram\'{e}r-Rao bound saturation tests, we find the best
 estimates of flux and centroid by optimizing the likelihood using the correct PSF.
 We expect this method to perform best in determing the centroids
 of stars, and deliver RMSE equal to Cram\'{e}r-Rao bound. 
\section{Centroiding methods}\label{sec:method}
\begin{description}
%\item{{\bf :N}} \quad \\ 
\item{{\bf Matched filter polynomial centroiding}} \quad Let us consider the case 
in which we have a good estimate of the pixel-convolved PSF at
the position of the faint star under consideration. 
We can smooth the image of the star, by correlating it with the 
full PSF $\mathcal{P}$ at the position of the star.
\begin{eqnarray}
Y^{(s)} &=& Y \star \mathcal{P}, \\
Y^{(s)}_{[i,j]} &=& \sum_{k,l}Y_{[i-k,j-l]}\mathcal{P}_{[k,l]},
\end{eqnarray}
where $Y$ is the image of the star, and $Y^{(s)}$ is sometimes called a matched filter. 
A matched filter is a method in which the data $Y$ is correlated (convolved in the 
case of symmetrical PSF) with the PSF $\mathcal{P}$. It is equivalent to optimizing the 
likelihood and therefore provides an optimal map 
where the peak of the map is the likely position of the 
point source (Lang \emph{et al.}, in preparation).

@@

When the curvature matrix $D$ has a zero (or very close to zero) deteminant,
centroid estimates obtained from equation (\ref{center}) can become arbitrarily 
large, which leads to catastrophic outliers. 
In order to tackle this issue, we add a soft regularization term
proportional to $\sigma^{2}$ to the diagonals of $D$ prior to inversion.
